###############################################################################################
Aim: Optimal thresholds for pan-cancer LLR6 and RF6 models
Description: To determine the optimal absolute and percentile thresholds for pan-cancer LLR6 and RF6 models (which maximizes Youden’s index on the training data)
###############################################################################################



# load required package
```{r}
library(data.table)
library(tidyverse)
library(ggridges)
library(readxl)
library(verification)
library(pROC)
library(scales)
library(stats)

```

# Set parameters and directories
```{r}
input_dir = "../02.Input/"
result_dir = "../03.Results/"

```

# load data 
```{r}
LLRmodelNA = 'LLR6' #  'LLR6'  'LLR5noTMB'  'LLR5noPSTH'

train_LLR6score_df = read_excel(paste0(result_dir,LLRmodelNA,'_Scaler(StandardScaler)_prediction.xlsx'), "0", col_names = TRUE)
train_RF6score_df = read_excel(paste0(result_dir,'RF6_Scaler(None)_prediction.xlsx'), "0", col_names = TRUE)

selected_cols = c("SAMPLE_ID", "CancerType", "Response")

train_info_df = read_excel(paste0(input_dir,'AllData.xlsx'), "Chowell_train")
train_info_df = train_info_df[selected_cols]

train_info_df = merge(x = train_info_df, y = train_LLR6score_df[c("SAMPLE_ID","y_pred")], by = "SAMPLE_ID")
train_info_df = merge(x = train_info_df, y = train_RF6score_df[c("SAMPLE_ID","y_pred")], by = "SAMPLE_ID")
colnames(train_info_df) = c(selected_cols, c("LLR6score","RF6score"))

CancerTypes = unique(train_info_df$CancerType)

```


# determine optimal LLR6score or RF6score percentile threshold by maximizing sensitivity + specificity
```{r}

LLR6_best_cutoff = 0
sum_score_max = 0
for (quant_cutoff in seq(0.1, 0.9, by=0.1)){  # 0.001
  train_df_temp = train_info_df %>% group_by(CancerType) %>% mutate(
         score_01 = ifelse(LLR6score >= quantile(LLR6score, quant_cutoff), 1, 0),
         )
  confusion <- confusionMatrix(as.factor(train_df_temp$score_01), as.factor(train_df_temp$Response))
  sensitivity <- confusion$byClass["Sensitivity"]
  specificity <- confusion$byClass["Specificity"]
  sum_score = sensitivity + specificity
  if (sum_score > sum_score_max){
    sum_score_max = sum_score
    LLR6_best_cutoff = quant_cutoff
  }
}
print(paste('LLR6 best percentile cutoff:', LLR6_best_cutoff))





RF6_best_cutoff = 0
sum_score_max = 0
for (quant_cutoff in seq(0.1, 0.9, by=0.1)){  # 0.001
  train_df_temp = train_info_df %>% group_by(CancerType) %>% mutate(
         score_01 = ifelse(RF6score >= quantile(RF6score, quant_cutoff), 1, 0),
         )
  confusion <- confusionMatrix(as.factor(train_df_temp$score_01), as.factor(train_df_temp$Response))
  sensitivity <- confusion$byClass["Sensitivity"]
  specificity <- confusion$byClass["Specificity"]
  sum_score = sensitivity + specificity
  if (sum_score > sum_score_max){
    sum_score_max = sum_score
    RF6_best_cutoff = quant_cutoff
  }
}
print(paste('RF6 best percentile cutoff:', RF6_best_cutoff))


```


# determine optimal LLR6score or RF6score absolute threshold by maximizing sensitivity + specificity
```{r}

LLR6_best_cutoff = 0
sum_score_max = 0
for (cutoff in seq(0.1, 0.9, by=0.001)){
  train_df_temp = train_info_df %>% group_by(CancerType) %>% mutate(
         score_01 = ifelse(LLR6score >= cutoff, 1, 0),
         )
  confusion <- confusionMatrix(as.factor(train_df_temp$score_01), as.factor(train_df_temp$Response))
  sensitivity <- confusion$byClass["Sensitivity"]
  specificity <- confusion$byClass["Specificity"]
  sum_score = sensitivity + specificity
  if (sum_score > sum_score_max){
    sum_score_max = sum_score
    LLR6_best_cutoff = cutoff
  }
  #print(paste(cutoff, sum_score))
}
print(paste('LLR6 best absolute cutoff:', LLR6_best_cutoff))




RF6_best_cutoff = 0
sum_score_max = 0
for (cutoff in seq(0.1, 0.9, by=0.001)){
  train_df_temp = train_info_df %>% group_by(CancerType) %>% mutate(
         score_01 = ifelse(RF6score >= cutoff, 1, 0),
         )
  confusion <- confusionMatrix(as.factor(train_df_temp$score_01), as.factor(train_df_temp$Response))
  sensitivity <- confusion$byClass["Sensitivity"]
  specificity <- confusion$byClass["Specificity"]
  sum_score = sensitivity + specificity
  if (sum_score > sum_score_max){
    sum_score_max = sum_score
    RF6_best_cutoff = cutoff
  }
}
print(paste('RF6 best absolute cutoff:', RF6_best_cutoff))


```
