###############################################################################################
Aim: Pan-cancer survival analysis in MSK non-ICB cohort
Description: To study LORIS predictive power in MSK non-ICB cohort (Fig. 5b,c,d)
(1) Multi-variable hazard ratio forest plot in individual cancers
(2) 0.5-,1-,2-,3-,4-,5-year PFS/OS in individual cancers
(3) K-M survival curves for 0-10%,10-20%,20-50%,50-80%,80-90%,90-100%-quantile groups of LORIS
###############################################################################################


# load required package
```{r}
library(data.table)
library(tidyverse)
library(survminer)
library(survival)
library(ggridges)
library(qvalue)
library(readxl)
library(verification)
library(pROC)
library(forestplot)
library(ggpubr)
library(pROC)

library(dplyr)
library(caret)
library(writexl)
library(viridis)

```

# Set parameters and directories
```{r}
input_dir = "../02.Input/"
result_dir = "../03.Results/"

LLRmodelNA = 'LLR6'

absoluteCutoff = 0.5

```

# N groups of K-M OS curves of LORIS
```{r}

modelNA = 'LLR6'
dataset = 'nonICB_test'
fontSize = 13


####### load ICB and nonICB data
ICB_test_df = read.csv(paste0(result_dir,'LLR6_predict_nonICB_vs_ICB_all_Dataset1.csv'))
ICB_test_df = ICB_test_df[c('CancerType', 'OS_Months', 'OS_Event','Age', modelNA)]

nonICB_test_df = read.csv(paste0(result_dir,'LLR6_predict_nonICB_vs_ICB_all_Dataset2.csv'))
nonICB_test_df = nonICB_test_df[c('CancerType', 'OS_Months', 'OS_Event','Age', modelNA)]


if (dataset == 'nonICB_test'){
  info_df = nonICB_test_df
}else if (dataset == 'ICB_test'){
  info_df = ICB_test_df
}

colnames(info_df) = c('CancerType', 'OS_Months', 'OS_Event','Age', 'LLRscore')

quantile_plot = info_df %>% group_by(CancerType) %>%
  mutate(Score_perct = ntile(LLRscore, n = n()) / n()) # ntile(Score, length(breaks)-1, breaks=breaks)

write.csv(quantile_plot, paste0(result_dir,"source_data_fig06b.csv"))


quantile_plot = quantile_plot %>% 
  mutate(Score_01 = ifelse(Score_perct > 0.9, 6, ifelse(Score_perct > 0.8, 5, ifelse(Score_perct > 0.5, 4, ifelse(Score_perct > 0.2, 3, ifelse(Score_perct > 0.1, 2, ifelse(Score_perct >= 0, 1, 0)))))))

survp=ggsurvplot(
  survfit(Surv(OS_Months, OS_Event) ~ Score_01, data = quantile_plot),
  data = quantile_plot,
  size = 1,                 # change line size
  palette = plasma(6),
  conf.int = FALSE,          # Add confidence interval
  pval = FALSE,              # Add p-value
  ylim=c(0,1),xlim=c(-1,55),
  xlab = "Time (months)", ylab="OS probability",
  break.time.by = 10,
  risk.table=TRUE,
  risk.table.height = 0.3, # Useful to change when you have multiple groups
  risk.table.col="black",
  risk.table.y.text = FALSE,
  tables.y.text = FALSE, 
  tables.theme = theme_cleantable(),
  legend.labs = c("0-10%", "10-20%", "20-50%", "50-80%", "80-90%", "90-100%"),
  legend.title="",
  legend = c(0.55, 0.25), # legend relative position
  font.main = c(fontSize),
  font.caption = c(fontSize),
  font.legend = c(fontSize),
  font.tickslab = c(fontSize),
  font.x = c(fontSize),
  font.y = c(fontSize),
 ggtheme = theme(
                 panel.background = element_rect(fill='transparent'), #transparent panel bg
                 plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
                 legend.background = element_rect(fill='transparent', color=NA), #transparent legend bg
                 legend.box.background = element_rect(fill='transparent', color=NA), #transparent legend panel
                 legend.key = element_rect(fill='transparent', color=NA),
                 plot.margin = unit(c(0.2, 0.2, 0.2, 0.2),"cm"),
                 panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                 panel.border = element_blank(),axis.line = element_line(colour = "black"),
                 axis.text.x = element_text(colour="black"),axis.text.y = element_text(colour="black")
                 ),  # top, right, bot, left
)

if (modelNA=="LLR6"){
  pdf(paste(result_dir,"MSK_",dataset,"_LLR6-ICB_6groups.pdf",sep = ""),width = 5, height = 6)
}else{
  pdf(paste(result_dir,"MSK_",dataset,"_LLR8-nonICB_6groups.pdf",sep = ""),width = 5, height = 6)
}

print(survp, newpage = FALSE)
dev.off()





##### p values between 5 different groups
quantile_plot_temp = quantile_plot[quantile_plot$Score_01==2 | quantile_plot$Score_01==1, ]

quantile_plot_temp = quantile_plot_temp %>% mutate(
         Score_01 = ifelse(Score_01 == 2, "high", "low"),
         )
quantile_plot_temp$Score_01 <- relevel(factor(quantile_plot_temp$Score_01), ref = "high")

scox = coxph(Surv(quantile_plot_temp$OS_Months,quantile_plot_temp$OS_Event)~Score_01, data=quantile_plot_temp)
scox_coef = summary(scox)$coefficients
HR_value = scox_coef[2] # hazard ratio
HR_CI = exp(confint(scox))
Z_value=scox_coef[4]
P_value= pnorm(Z_value, lower.tail = FALSE) # to test if higher score has better survival
print(paste0('HR_value: ', round(HR_value,2), '.  P_value: ', P_value))
```

# Binarize LLR6 and TMB values
```{r}

info_df = nonICB_test_df

info_df = info_df %>% group_by(CancerType) %>% mutate(
         LLRscore_01 = ifelse(LLR6 >= absoluteCutoff, "high", "low"),
         LLRscore_01_2 = ifelse(LLR6 >= quantile(LLR6, 0.5), "high", "low"),
         )
info_df$LLRscore_01 <- relevel(factor(info_df$LLRscore_01), ref = "low")
info_df$LLRscore_01_2 <- relevel(factor(info_df$LLRscore_01_2), ref = "low")

```


# Univariate/Multivariable hazard ratio Forest plot (OS/PFS, continuous/binary)  HR_each_cancer
```{r}

CancerTypes = unique(info_df$CancerType)

survival_type = 'OS' # 'PFS'  'OS'
HR_type = 'Multi' # Multi Uni
Var_type = 'binaryPercent' # continuous  binaryAbsolute binaryPercent
testVAR0 = 'LLRscore' # RF6score LLRscore TMB
if (Var_type == 'binaryAbsolute'){
  testVAR = paste0(testVAR0, '_01')
}else if (Var_type == 'binaryPercent'){
  testVAR = paste0(testVAR0, '_01_2')
}
testVAR_HR = testVAR
if (grepl("binary", Var_type)){
  testVAR_HR = paste0(testVAR,'high')
}
if (HR_type == 'Multi'){
  var_names = c(testVAR,'Age')
}else{
  var_names = c(testVAR)
}


HR_pancancer = data.frame()
cancerType = 'All'
data_for_HR = info_df
formula_obj <- as.formula(paste(paste0("Surv(",survival_type,"_Months, ",survival_type,"_Event) ~CancerType+"), paste(var_names, collapse = " + ")))
model = coxph(formula_obj, data = data_for_HR)

HR_ct = exp(model$coefficients[[testVAR_HR]])
HR_ct_025 = exp(confint(model))[testVAR_HR,1]
HR_ct_975 = exp(confint(model))[testVAR_HR,2]
p_ct = summary(model)$coefficients[testVAR_HR,5]

HR_pancancer = rbind(HR_pancancer,data.frame(HR_ct,HR_ct_025,HR_ct_975,p_ct,dim(data_for_HR)[1],cancerType))
colnames(HR_pancancer) = c(paste0('HR_',testVAR),paste0('HR_',testVAR,'025'),paste0('HR_',testVAR,'975'),paste0('p-val_',testVAR),'No. of patients', 'CancerType')

HR_each_cancer = data.frame()
formula_obj <- as.formula(paste(paste0("Surv(",survival_type,"_Months, ",survival_type,"_Event) ~"), paste(var_names, collapse = " + ")))
for (ct in CancerTypes){
  cancerType = ct
  model = coxph(formula_obj, data = data_for_HR[data_for_HR$CancerType==ct,])
  HR_ct = exp(model$coefficients[[testVAR_HR]])
  HR_ct_025 = exp(confint(model))[testVAR_HR,1]
  HR_ct_975 = exp(confint(model))[testVAR_HR,2]
  p_ct = summary(model)$coefficients[testVAR_HR,5]
  HR_each_cancer = rbind(HR_each_cancer,data.frame(HR_ct,HR_ct_025,HR_ct_975,p_ct,dim(data_for_HR[data_for_HR$CancerType==ct,])[1],cancerType))
}
colnames(HR_each_cancer) = c(paste0('HR_',testVAR),paste0('HR_',testVAR,'025'),paste0('HR_',testVAR,'975'),paste0('p-val_',testVAR),'No. of patients', 'CancerType')


command_str <- paste0("HR_each_cancer = arrange(HR_each_cancer, desc(HR_",testVAR,"))") # desc(is.na(HR_",testVAR,")),
eval(parse(text = command_str))
cancerOrder = HR_each_cancer$CancerType

HR_each_cancer = rbind(HR_each_cancer,HR_pancancer)

write.csv(HR_each_cancer, paste0(result_dir,"source_data_fig06c.csv"))



plot_data <- tibble::tibble(mean  = HR_each_cancer[[paste0('HR_',testVAR)]],
                            lower = HR_each_cancer[[paste0('HR_',testVAR,'025')]],
                            upper = HR_each_cancer[[paste0('HR_',testVAR,'975')]],
                            cancer = HR_each_cancer$CancerType,
                            pval = HR_each_cancer[[paste0('p-val_',testVAR)]],
                            OR = round(HR_each_cancer[[paste0('HR_',testVAR)]], 2),
                            size = HR_each_cancer$`No. of patients`)

options(scipen = 999)
P_value_raw = c(plot_data[['pval']], HR_pancancer[[paste0('p-val_',testVAR)]])


pval_vec = vector("character", length(P_value_raw))
for (i in 1:length(P_value_raw)){
  pval = P_value_raw[i]
  if (pval>=0.1){
    pval_vec[i] = as.character(round(pval,2))
  }else{
    pval_vec[i] = format(pval, scientific = TRUE, digits = 2)  # Format for scientific notation
  }
}

plot_data$pval = pval_vec[1:length(pval_vec)-1]

if (testVAR=='LLRscore'){ # testVAR=='LLRscore'
  xmin_lim = 0
  xmax_lim = 10
  breaks_x = c(0,1,10)
}else if (testVAR=="LLRscore_01" | testVAR=="LLRscore_01_2"){ # testVAR=='LLRscore'
  xmin_lim = 0
  xmax_lim = 4
  breaks_x = c(0,1,2,3,4)
}else if (testVAR=='TMB'){ # testVAR=='TMB'
  xmin_lim = 0
  xmax_lim = 2
  breaks_x = c(0,1,2)
}else if (testVAR=="TMB_01" | testVAR=="TMB_01_2"){ # testVAR=='LLRscore'
  xmin_lim = 0
  xmax_lim = 4
  breaks_x = c(0,1,2,3,4)
}
labels_x = breaks_x


fontSize = 1.2

pdf_file <- paste0(result_dir,paste0("MSK_nonICB_Forest_",survival_type,"_",HR_type,"_",Var_type,"_",testVAR0,".pdf"))
fig_width = 5
fig_height = 5.5
pdf(pdf_file, onefile=FALSE, width = fig_width,height=fig_height)


plot_data %>%
  forestplot(labeltext = c("cancer", "size", "pval"),
             graph.pos = 3,
             boxsize = 0.35,
             #line.margin = .30, # We need to add this to avoid crowding
             vertices = TRUE,
             clip = c(xmin_lim, xmax_lim),
             xlog = FALSE,
             zero = 1,
             #txt_gp = fpTxtGp(cex=0.75),
             txt_gp = fpTxtGp(ticks=gpar(cex=fontSize),xlab=gpar(cex=fontSize),label=gpar(cex=fontSize),legend=gpar(cex=fontSize),title=gpar(cex=fontSize)), # label, summary, xlab, title, ticks legend
             xlab = "Hazard ratio",
             xticks = breaks_x,
             #lineheight=unit(1,'cm'),
             graphwidth = unit(3, "cm"),
             lineheight = unit(2, "cm")
             ) %>%
  fp_set_style(box = "black",
               line = "black",
               summary = "black") %>%
  fp_add_header(cancer = c("Cancer type")|> fp_txt_plain() |> fp_align_center(),
                size = c("Samples")|> fp_txt_plain() |> fp_align_center(), # "No. of patients"
                pval = c("P-value")|> fp_txt_plain() |> fp_align_center())

dev.off()

```


# X-year OS comparison for score-low VS score-high patients (X = 0.5,1,2,3,4,5 years)
```{r}
testVAR = 'LLRscore_01_2' # RF6score_01 LLRscore_01 TMB_01 RF6score_01_2 LLRscore_01_2 TMB_01_2

if (grepl('_01_2', testVAR)){
  testVAR_NA = paste0(substr(testVAR, 1, nchar(testVAR) - 5),'_PercentileCutoff')
}else{
  testVAR_NA = paste0(substr(testVAR, 1, nchar(testVAR) - 3),'_AbsoluteCutoff')
}

OSrate_score = data.frame()

for (os_time in c(6,12,24,36,48,60)){
  month_OS = os_time
  for (ct in CancerTypes){
    
    data_ct = info_df[info_df['CancerType']==ct,]
    cancerData=data.frame(data_ct[testVAR],data_ct$OS_Months,data_ct$OS_Event)
    colnames(cancerData) = c("Score", "OS_Months", "OS_Event")
    sfit = survfit(Surv(OS_Months,OS_Event) ~ Score, data=cancerData)
    OS_rates = summary(sfit,time=c(month_OS),extend=TRUE)$surv
    OSrate_score = rbind(OSrate_score,data.frame(OS_rates[1],OS_rates[2],month_OS,row.names = ct))
  }
  cancerData=data.frame(info_df[testVAR],info_df$OS_Months,info_df$OS_Event)
  colnames(cancerData) = c("Score", "OS_Months", "OS_Event")
  sfit = survfit(Surv(OS_Months,OS_Event) ~ Score, data=cancerData)
  OS_rates = summary(sfit,time=c(month_OS),extend=TRUE)$surv
  OSrate_score = rbind(OSrate_score, data.frame(OS_rates[1],OS_rates[2],month_OS,row.names = 'All'))
}
colnames(OSrate_score) = c('OSR_L','OSR_H','Time')

OSrate_score['CANCER_TYPE'] = rownames(OSrate_score)

write.csv(OSrate_score, paste0(result_dir,"source_data_fig06d.csv"))



################ 6-month OSR paired plot
OSrate_score_6month = OSrate_score[OSrate_score$Time==6,]
ggpaired(OSrate_score_6month %>% rename(`High` = OSR_H,
                                `Low` = OSR_L), 
         cond1 = "High", cond2 = "Low", line.color = "CANCER_TYPE", line.size = 1,
         fill = "gray65",
         palette = plasma(17), xlab = "Half year", legend.title = "Cancer type", ylab = "OS probability") + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust = 0.5)) + # right angle = 30
  ylim(0, 1) 
ggsave(paste0(result_dir,paste0("MSK_nonICB_OS_6month_Comparison_",testVAR_NA,".pdf")), width = 1.8*1.2, height = 2.4*1.2) # width = 2.3, height = 2.7
print(paste("Delta 6-month OS rate:",median(OSrate_score_6month$OSR_H, na.rm = TRUE) - median(OSrate_score_6month$OSR_L, na.rm = TRUE),sep=" "))
stat_result = wilcox.test(OSrate_score_6month$OSR_H, OSrate_score_6month$OSR_L, paired = TRUE, alternative = "two.sided", na.rm = TRUE)
print(paste0('6-month Wilcoxon p-value: ',stat_result$p.value))

################ 12-month OSR paired plot
OSrate_score_12month = OSrate_score[OSrate_score$Time==12,]
ggpaired(OSrate_score_12month %>% rename(`High` = OSR_H,
                                `Low` = OSR_L), 
         cond1 = "High", cond2 = "Low", line.color = "CANCER_TYPE", line.size = 1,
         fill = "gray65",
         palette = plasma(17), xlab = "One year", legend.title = "Cancer type", ylab = "OS probability") + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust = 0.5)) + # right
  ylim(0, 1) 
ggsave(paste0(result_dir,paste0("MSK_nonICB_OS_12month_Comparison_",testVAR_NA,".pdf")), width = 1.8*1.2, height = 2.4*1.2) # width = 2.3, height = 2.7
print(paste("Delta 12-month OS rate:",median(OSrate_score_12month$OSR_H, na.rm = TRUE) - median(OSrate_score_12month$OSR_L, na.rm = TRUE),sep=" "))
stat_result = wilcox.test(OSrate_score_12month$OSR_H, OSrate_score_12month$OSR_L, paired = TRUE, alternative = "two.sided", na.rm = TRUE)
print(paste0('12-month Wilcoxon p-value: ',stat_result$p.value))

################ 24-month OSR paired plot
OSrate_score_24month = OSrate_score[OSrate_score$Time==24,]
ggpaired(OSrate_score_24month %>% rename(`High` = OSR_H,
                                `Low` = OSR_L), 
         cond1 = "High", cond2 = "Low", line.color = "CANCER_TYPE", line.size = 1,
         fill = "gray65",
         palette = plasma(17), xlab = "Two years", legend.title = "Cancer type", ylab = "OS probability") + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust = 0.5)) + # right
  ylim(0, 1) 
ggsave(paste0(result_dir,paste0("MSK_nonICB_OS_24month_Comparison_",testVAR_NA,".pdf")), width = 1.8*1.2, height = 2.4*1.2) # width = 2.3, height = 2.7
print(paste("Delta 24-month OS rate:",median(OSrate_score_24month$OSR_H, na.rm = TRUE) - median(OSrate_score_24month$OSR_L, na.rm = TRUE),sep=" "))
stat_result = wilcox.test(OSrate_score_24month$OSR_H, OSrate_score_24month$OSR_L, paired = TRUE, alternative = "two.sided", na.rm = TRUE)
print(paste0('24-month Wilcoxon p-value: ',stat_result$p.value))

################ 36-month OSR paired plot
OSrate_score_36month = OSrate_score[OSrate_score$Time==36,]
ggpaired(OSrate_score_36month %>% rename(`High` = OSR_H,
                                `Low` = OSR_L), 
         cond1 = "High", cond2 = "Low", line.color = "CANCER_TYPE", line.size = 1,
         fill = "gray65",
         palette = plasma(17), xlab = "Three years", legend.title = "Cancer type", ylab = "OS probability") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust = 0.5)) + # right
  ylim(0, 1) 
ggsave(paste0(result_dir,paste0("MSK_nonICB_OS_36month_Comparison_",testVAR_NA,".pdf")), width = 1.8*1.2, height = 2.4*1.2) # width = 2.3, height = 2.7
print(paste("Delta 36-month OS rate:",median(OSrate_score_36month$OSR_H, na.rm = TRUE) - median(OSrate_score_36month$OSR_L, na.rm = TRUE),sep=" "))
stat_result = wilcox.test(OSrate_score_36month$OSR_H, OSrate_score_36month$OSR_L, paired = TRUE, alternative = "two.sided", na.rm = TRUE)
print(paste0('36-month Wilcoxon p-value: ',stat_result$p.value))

################ 48-month OSR paired plot
OSrate_score_48month = OSrate_score[OSrate_score$Time==48,]
ggpaired(OSrate_score_48month %>% rename(`High` = OSR_H,
                                `Low` = OSR_L), 
         cond1 = "High", cond2 = "Low", line.color = "CANCER_TYPE", line.size = 1,
         fill = "gray65",
         palette = plasma(17), xlab = "Four years", legend.title = "Cancer type", ylab = "OS probability") + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust = 0.5)) + # right
  ylim(0, 1) 
ggsave(paste0(result_dir,paste0("MSK_nonICB_OS_48month_Comparison_",testVAR_NA,".pdf")), width = 1.8*1.2, height = 2.4*1.2) # width = 2.3, height = 2.7
print(paste("Delta 48-month OS rate:",median(OSrate_score_48month$OSR_H, na.rm = TRUE) - median(OSrate_score_48month$OSR_L, na.rm = TRUE),sep=" "))
stat_result = wilcox.test(OSrate_score_48month$OSR_H, OSrate_score_48month$OSR_L, paired = TRUE, alternative = "two.sided", na.rm = TRUE)
print(paste0('48-month Wilcoxon p-value: ',stat_result$p.value))

################ 60-month OSR paired plot
OSrate_score_60month = OSrate_score[OSrate_score$Time==60,]
ggpaired(OSrate_score_60month %>% rename(`High` = OSR_H,
                                `Low` = OSR_L), 
         cond1 = "High", cond2 = "Low", line.color = "CANCER_TYPE", line.size = 1,
         fill = "gray65",
         palette = plasma(17), xlab = "Five years", legend.title = "Cancer type", ylab = "OS probability") + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust = 0.5)) + # right
  ylim(0, 1) 
ggsave(paste0(result_dir,paste0("MSK_nonICB_OS_60month_Comparison_",testVAR_NA,".pdf")), width = 1.8*1.2, height = 2.4*1.2) # width = 2.3, height = 2.7
print(paste("Delta 60-month OS rate:",median(OSrate_score_60month$OSR_H, na.rm = TRUE) - median(OSrate_score_60month$OSR_L, na.rm = TRUE),sep=" "))
stat_result = wilcox.test(OSrate_score_60month$OSR_H, OSrate_score_60month$OSR_L, paired = TRUE, alternative = "two.sided", na.rm = TRUE)
print(paste0('60-month Wilcoxon p-value: ',stat_result$p.value))

```