###############################################################################################
Aim: Kaplan-Meier analysis of PFS/OS in individual cancers
Description: To plot Kaplan-Meier curves of PFS/OS in individual cancers of merged MSK cohorts (Supplementary Figs. 4,5)
###############################################################################################



# load required package
```{r}
library(data.table)
library(tidyverse)
library(survminer)
library(survival)
library(ggridges)
library(qvalue)
library(readxl)
library(verification)
library(pROC)
library(forestplot)

library(viridis)
library(patchwork)
library(gridExtra)

```

# Set parameters and directories
```{r}
input_dir = "../02.Input/"
result_dir = "../03.Results/"

LLRmodelNA = 'LLR6'
LLR6_percentCutoff = 0.5
TMB_percentCutoff = 0.8

```

# load data 
```{r}
train_LLR6score_df = read_excel(paste0(result_dir,'PanCancer_all_LLR6_Scaler(StandardScaler)_prediction.xlsx'), "0", col_names = TRUE)

test1_LLR6score_df = read_excel(paste0(result_dir,'PanCancer_all_LLR6_Scaler(StandardScaler)_prediction.xlsx'), "1", col_names = TRUE)
test2_LLR6score_df = read_excel(paste0(result_dir,'PanCancer_all_LLR6_Scaler(StandardScaler)_prediction.xlsx'), "2")
test3_LLR6score_df = read_excel(paste0(result_dir,'PanCancer_all_LLR6_Scaler(StandardScaler)_prediction.xlsx'), "3")
LLR6score_df = rbind(train_LLR6score_df,test1_LLR6score_df,test2_LLR6score_df,test3_LLR6score_df)

selected_cols = c("SAMPLE_ID","Year of ICI treatment start", "TMB", "Chemo_before_IO", "CancerType", "Age", "Drug", "Sex", "PFS_Event", "PFS_Months", "OS_Event", "OS_Months")

train_info_df = read_excel(paste0(input_dir,'features_phenotype_allDatasets.xlsx'), "Chowell2015-2017")
train_info_df = train_info_df[selected_cols]
test1_info_df = read_excel(paste0(input_dir,'features_phenotype_allDatasets.xlsx'), "Chowell2018")
test1_info_df = test1_info_df[selected_cols]
test2_info_df = read_excel(paste0(input_dir,'features_phenotype_allDatasets.xlsx'), "Morris_new")
test2_info_df = test2_info_df[selected_cols]
test3_info_df = read_excel(paste0(input_dir,'features_phenotype_allDatasets.xlsx'), "Morris_new2")
test3_info_df = test3_info_df[selected_cols]

renamed_cols = c("SAMPLE_ID","Year_IO", "TMB", "Chemo_before_IO", "CancerType", "Age", "Drug", "Sex", "PFS_Event", "PFS_Months", "OS_Event", "OS_Months")

info_df = rbind(train_info_df,test1_info_df,test2_info_df,test3_info_df)
info_df = merge(x = info_df, y = LLR6score_df[c("SAMPLE_ID","y_pred")], by = "SAMPLE_ID")
colnames(info_df) = c(renamed_cols, c("LLR6score"))



CancerTypes = unique(info_df$CancerType)

```



# K-M PFS analysis for individual cancer types (merged MSK cohort)
```{r}

stratify_var = 'LORIS-'

info_df = info_df %>% group_by(CancerType) %>% mutate(
         LLR6score_01 = ifelse(LLR6score >= quantile(LLR6score,LLR6_percentCutoff), "LLR6high", "LLR6low"),
         TMB_01 = ifelse(TMB >= quantile(TMB,TMB_percentCutoff), "TMBhigh", "TMBlow"),
         )

data_plot = info_df
data_plot[['LLR6score_01']] <- relevel(factor(data_plot[['LLR6score_01']]), ref = "LLR6high")

for (ct in CancerTypes){
  data_temp = data_plot[data_plot$CancerType==ct,]
  cancerData=data.frame(data_temp['LLR6score_01'],data_temp$PFS_Months,data_temp$PFS_Event)
  colnames(cancerData) = c("Score", "PFS_Months", "PFS_Event")
  sfit = survfit(Surv(data_temp$PFS_Months,data_temp$PFS_Event) ~ Score, data=cancerData)
  scox = coxph(Surv(data_temp$PFS_Months,data_temp$PFS_Event)~Score, data=cancerData)
  scox_coef = summary(scox)$coefficients
  HR_value = scox_coef[2] # hazard ratio
  Z_value=scox_coef[4]
  P_value=scox_coef[5]
  HR_CI = exp(confint(scox))
  print(paste(c('PFS', ct, round(P_value,5), round(HR_value,2)), collapse= " "))
  ##### plot
  fontSize = 13
  survp=ggsurvplot(
    sfit,
    data = cancerData,
    size = 1,                 # change line size
    palette = c("#377EB8", "#FF7F0E"), #viridis_pal()(2), #plasma(2),
    conf.int = F,          # Add confidence interval
    pval = FALSE,              # Add p-value
    ylim=c(0,1),
    xlab = "Time (months)", ylab="PFS probability",# Time in months
    break.time.by = 10,
    risk.table=TRUE,
    risk.table.height = 0.25, # Useful to change when you have multiple groups
    risk.table.pos="out",
    risk.table.col="black",
    risk.table.y.text = FALSE,
    tables.y.text = FALSE, 
    tables.theme = theme_cleantable(),
    legend.labs =c(paste0(stratify_var," high"),paste0(stratify_var," low")),    # Change legend labels
    legend.title="",
    legend = c(2.65, 0.9), # legend relative position
    font.main = c(fontSize),
    font.caption = c(fontSize),
    font.legend = c(fontSize),
    font.tickslab = c(fontSize),
    font.x = c(fontSize),
    font.y = c(fontSize),
    ggtheme = theme(legend.background = element_rect(fill = NA, color=NA),legend.key = element_rect(fill = NA, color=NA),
                   plot.margin = unit(c(0.2, 0.2, 0, 0.2),"cm"),
                   panel.background = element_rect(fill = "white"),
                   panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                   panel.border = element_blank(),axis.line = element_line(colour = "black"),
                   axis.text.x = element_text(colour="black"),axis.text.y = element_text(colour="black")),  # top, right, bot, left
  )+ 
    guides(colour = guide_legend(nrow = 1)) # legend in rows
  survp$plot = survp$plot+ 
              ggplot2::annotate("text", x=0, y=0.2, label=paste0('p = ',sprintf('%.1e', P_value), '\n', 'HR',' = ',round(HR_value,2),' (', round(HR_CI[1],2),'-',round(HR_CI[2],2),')'),size = 5, hjust = 0)
  title_str = paste0('MSK_allMerged_',ct,'_PFS_LLR6score') # round(P_value,3)
  pdf(paste(result_dir,title_str,".pdf",sep = ""),width=3, height=3)
  print(survp, newpage = FALSE)
  dev.off()
}

```

# K-M OS analysis for individual cancer types (merged MSK cohort)
```{r}

stratify_var = 'LORIS-'

info_df = info_df %>% group_by(CancerType) %>% mutate(
         LLR6score_01 = ifelse(LLR6score >= quantile(LLR6score,LLR6_percentCutoff), "LLR6high", "LLR6low"),
         TMB_01 = ifelse(TMB >= quantile(TMB,TMB_percentCutoff), "TMBhigh", "TMBlow"),
         )

data_plot = info_df
data_plot[['LLR6score_01']] <- relevel(factor(data_plot[['LLR6score_01']]), ref = "LLR6high")

write.csv(data_plot, paste0(result_dir,"source_data_fig04.csv"))


for (ct in CancerTypes){
  data_temp = data_plot[data_plot$CancerType==ct,]
  cancerData=data.frame(data_temp['LLR6score_01'],data_temp$OS_Months,data_temp$OS_Event)
  colnames(cancerData) = c("Score", "OS_Months", "OS_Event")
  sfit = survfit(Surv(data_temp$OS_Months,data_temp$OS_Event) ~ Score, data=cancerData)
  scox = coxph(Surv(data_temp$OS_Months,data_temp$OS_Event)~Score, data=cancerData)
  scox_coef = summary(scox)$coefficients
  HR_value = scox_coef[2] # hazard ratio
  Z_value=scox_coef[4]
  P_value=scox_coef[5]
  HR_CI = exp(confint(scox))
  print(paste(c('OS', ct, round(P_value,5), round(HR_value,2)), collapse= " "))
  ##### plot
  fontSize = 13
  survp=ggsurvplot(
    sfit,
    data = cancerData,
    size = 1,                 # change line size
    palette = c("#377EB8", "#FF7F0E"), #viridis_pal()(2), #plasma(2),
    conf.int = F,          # Add confidence interval
    pval = FALSE,              # Add p-value
    ylim=c(0,1),
    xlab = "Time (months)", ylab="OS probability",# Time in months
    break.time.by = 10,
    risk.table=TRUE,
    risk.table.height = 0.25, # Useful to change when you have multiple groups
    risk.table.pos="out",
    risk.table.col="black",
    risk.table.y.text = FALSE,
    tables.y.text = FALSE, 
    tables.theme = theme_cleantable(),
    legend.labs =c(paste0(stratify_var," high"),paste0(stratify_var," low")),    # Change legend labels
    legend.title="",
    legend = c(0.65, 0.9), # legend relative position
    font.main = c(fontSize),
    font.caption = c(fontSize),
    font.legend = c(fontSize),
    font.tickslab = c(fontSize),
    font.x = c(fontSize),
    font.y = c(fontSize),
    ggtheme = theme(legend.background = element_rect(fill = NA, color=NA),
                    legend.key = element_rect(fill = NA, color=NA),
                   plot.margin = unit(c(0.2, 0.2, 0, 0.2),"cm"),
                   panel.background = element_rect(fill = "white"),
                   panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                   panel.border = element_blank(),axis.line = element_line(colour = "black"),
                   axis.text.x = element_text(colour="black"),axis.text.y = element_text(colour="black")),  # top, right, bot, left
  )+ 
    guides(colour = guide_legend(ncol = 1)) # nrow = 1
  survp$plot = survp$plot+ 
              ggplot2::annotate("text", x=0, y=0.2, label=paste0('p = ',sprintf('%.1e', P_value), '\n', 'HR',' = ',round(HR_value,2),' (', round(HR_CI[1],2),'-',round(HR_CI[2],2),')'),size = 5, hjust = 0)
  title_str = paste0('MSK_allMerged_',ct,'_OS_LLR6score') # round(P_value,3)
  pdf(paste(result_dir,title_str,".pdf",sep = ""),width=3, height=3)
  print(survp, newpage = FALSE)
  dev.off()
}

```


