###############################################################################################
Aim: Predict survival of NSCLC patients using different models
Description: calculate multi-variable hazard ratios for different cohorts using LORIS, PDL1 & TMB
LORIS has been calculated using NSCLC-specific LLR6, pan-cancer LLR6, and NSCLC-specific LLR2 models respectively
(Fig. 6d,e; Extended Data Fig. 6b,c,e,f)
###############################################################################################


# load required package
```{r}
library(data.table)
library(tidyverse)
library(survminer)
library(survival)
library(ggridges)
library(qvalue)
library(readxl)
library(verification)
library(pROC)
library(forestplot)
```

# Set parameters and directories
```{r}
input_dir = "../02.Input/"
result_dir = "../03.Results/"

LLRmodelNA = 'LLR2' # LLR6   'LLR2'   'LLR6Pan'

if (LLRmodelNA=='LLR6'){
  absoluteCutoff = 0.44
}else if (LLRmodelNA=='LLR2'){
  absoluteCutoff = 0.46
}else if (LLRmodelNA=='LLR6Pan'){
  absoluteCutoff = 0.5
}


```

# load data 
```{r}
train_LLR6score_df = read_excel(paste0(result_dir,'NSCLC_',LLRmodelNA,'_Scaler(StandardScaler)_prediction.xlsx'), "0", col_names = TRUE)
test1_LLR6score_df = read_excel(paste0(result_dir,'NSCLC_',LLRmodelNA,'_Scaler(StandardScaler)_prediction.xlsx'), "1", col_names = TRUE)
test2_LLR6score_df = read_excel(paste0(result_dir,'NSCLC_',LLRmodelNA,'_Scaler(StandardScaler)_prediction.xlsx'), "2")
test3_LLR6score_df = read_excel(paste0(result_dir,'NSCLC_',LLRmodelNA,'_Scaler(StandardScaler)_prediction.xlsx'), "3")
test4_LLR6score_df = read_excel(paste0(result_dir,'NSCLC_',LLRmodelNA,'_Scaler(StandardScaler)_prediction.xlsx'), "4")

train_LLR6score_df$DataSet <- "Chowell et al."
test1_LLR6score_df$DataSet <- "MSK1"
test2_LLR6score_df$DataSet <- "Shim et al."
test3_LLR6score_df$DataSet <- "Vanguri et al."
test4_LLR6score_df$DataSet <- "Ravi et al."

all_LLR6score_df = rbind(train_LLR6score_df,test1_LLR6score_df,test2_LLR6score_df,test3_LLR6score_df,test4_LLR6score_df)

selected_cols = c("SAMPLE_ID","TMB","PDL1_TPS(%)","Systemic_therapy_history","Age","Drug","Sex","PFS_Event","PFS_Months","OS_Event","OS_Months")

train1_info_df = read_excel(paste0(input_dir,'AllData.xlsx'), "Chowell_train")
train1_info_df = train1_info_df[train1_info_df$CancerType=='NSCLC',selected_cols]
train2_info_df = read_excel(paste0(input_dir,'AllData.xlsx'), "Chowell_test")
train2_info_df = train2_info_df[train2_info_df$CancerType=='NSCLC',selected_cols]
train_info_df  = rbind(train1_info_df,train2_info_df)
test1_info_df = read_excel(paste0(input_dir,'AllData.xlsx'), "MSK1")
test1_info_df = test1_info_df[test1_info_df$CancerType=='NSCLC',selected_cols]
test2_info_df = read_excel(paste0(input_dir,'AllData.xlsx'), "Shim_NSCLC")
test2_info_df = test2_info_df[selected_cols]
test3_info_df = read_excel(paste0(input_dir,'AllData.xlsx'), "Vanguri_NSCLC")
test3_info_df = test3_info_df[selected_cols]
test4_info_df = read_excel(paste0(input_dir,'AllData.xlsx'), "Ravi_NSCLC")
test4_info_df = test4_info_df[selected_cols]

info_df = rbind(train_info_df,test1_info_df,test2_info_df,test3_info_df,test4_info_df)

info_df = merge(x = info_df, y = all_LLR6score_df[c("SAMPLE_ID","y_pred", "DataSet")], by = "SAMPLE_ID")
colnames(info_df) = c("SAMPLE_ID","TMB","PDL1","Systemic_therapy_history","Age","Drug","Sex","PFS_Event","PFS_Months","OS_Event","OS_Months", "LLRscore", "DataSet")

DataSets = c('Chowell et al.', 'MSK1', 'Shim et al.', 'Vanguri et al.', 'Ravi et al.')
```


# Binarize LLR6, PD-L1 and TMB values
```{r}
info_df = info_df %>% group_by(DataSet) %>% mutate(
         LLRscore_01 = ifelse(LLRscore >= absoluteCutoff, "high", "low"),
         PDL1_01 = ifelse(PDL1 >= 50, "high", "low"),
         TMB_01 = ifelse(TMB >= 10, "high", "low"),
         LLRscore_01_2 = ifelse(LLRscore >= quantile(LLRscore, 0.5), "high", "low"),
         PDL1_01_2 = ifelse(PDL1 >= quantile(PDL1, 0.8), "high", "low"),
         TMB_01_2 = ifelse(TMB >= quantile(TMB, 0.8), "high", "low"),
         )
info_df$LLRscore_01 <- relevel(factor(info_df$LLRscore_01), ref = "low")
info_df$PDL1_01 <- relevel(factor(info_df$PDL1_01), ref = "low")
info_df$TMB_01 <- relevel(factor(info_df$TMB_01), ref = "low")
info_df$LLRscore_01_2 <- relevel(factor(info_df$LLRscore_01_2), ref = "low")
info_df$PDL1_01_2 <- relevel(factor(info_df$PDL1_01_2), ref = "low")
info_df$TMB_01_2 <- relevel(factor(info_df$TMB_01_2), ref = "low")

```

# Univariate/Multivariable hazard ratio Forest plot (OS/PFS, continuous/binary)  HR_each_cancer
```{r}
survival_type = 'OS' # 'PFS'  'OS'
HR_type = 'Multi' # Multi Uni
Var_type = 'binaryAbsolute' # continuous  binaryAbsolute binaryPercent
testVAR0 = 'LLRscore' # LLRscore TMB PDL1
testVAR_name = 'LORIS' # LORIS    PDL1 TPS   TMB
if (Var_type == 'binaryAbsolute'){
  testVAR = paste0(testVAR0, '_01')
}else if (Var_type == 'binaryPercent'){
  testVAR = paste0(testVAR0, '_01_2')
}
testVAR_HR = testVAR
if (grepl("binary", Var_type)){
  testVAR_HR = paste0(testVAR,'high')
}
if (HR_type == 'Multi'){
  var_names = c(testVAR,'Drug', 'Age','Sex')
}else{
  var_names = c(testVAR)
}


HR_pancohort = data.frame()
DataSet = 'All'
data_for_HR = info_df
formula_obj <- as.formula(paste(paste0("Surv(",survival_type,"_Months, ",survival_type,"_Event) ~"), paste(var_names, collapse = " + ")))
model = coxph(formula_obj, data = data_for_HR)

HR_ct = exp(model$coefficients[[testVAR_HR]])
HR_ct_025 = exp(confint(model))[testVAR_HR,1]
HR_ct_975 = exp(confint(model))[testVAR_HR,2]
p_ct = summary(model)$coefficients[testVAR_HR,5]

HR_pancohort = rbind(HR_pancohort,data.frame(HR_ct,HR_ct_025,HR_ct_975,p_ct,dim(data_for_HR)[1],DataSet))
colnames(HR_pancohort) = c(paste0('HR_',testVAR),paste0('HR_',testVAR,'025'),paste0('HR_',testVAR,'975'),paste0('p-val_',testVAR),'No. of patients', 'DataSet')

HR_each_cancer = data.frame()
for (ct in DataSets){
  DataSet = ct
  model = coxph(formula_obj, data = data_for_HR[data_for_HR$DataSet==ct,])
  HR_ct = exp(model$coefficients[[testVAR_HR]])
  HR_ct_025 = exp(confint(model))[testVAR_HR,1]
  HR_ct_975 = exp(confint(model))[testVAR_HR,2]
  p_ct = summary(model)$coefficients[testVAR_HR,5]
  HR_each_cancer = rbind(HR_each_cancer,data.frame(HR_ct,HR_ct_025,HR_ct_975,p_ct,dim(data_for_HR[data_for_HR$DataSet==ct,])[1],DataSet))
}
colnames(HR_each_cancer) = c(paste0('HR_',testVAR),paste0('HR_',testVAR,'025'),paste0('HR_',testVAR,'975'),paste0('p-val_',testVAR),'No. of patients', 'DataSet')



if (testVAR0 == 'LLRscore'){
  command_str <- paste0("HR_each_cancer = arrange(HR_each_cancer, desc(HR_",testVAR,"))") # desc(is.na(HR_",testVAR,")),
  eval(parse(text = command_str))
  cancerOrder = HR_each_cancer$DataSet
}else{
  rownames(HR_each_cancer) = HR_each_cancer$DataSet
  HR_each_cancer = HR_each_cancer[cancerOrder,]
}


HR_each_cancer = rbind(HR_each_cancer,HR_pancohort)



plot_data <- tibble::tibble(mean  = HR_each_cancer[[paste0('HR_',testVAR)]],
                            lower = HR_each_cancer[[paste0('HR_',testVAR,'025')]],
                            upper = HR_each_cancer[[paste0('HR_',testVAR,'975')]],
                            dataset = HR_each_cancer$DataSet,
                            pval = HR_each_cancer[[paste0('p-val_',testVAR)]],
                            OR = round(HR_each_cancer[[paste0('HR_',testVAR)]], 2),
                            size = HR_each_cancer$`No. of patients`)

options(scipen = 999)
P_value_raw = c(plot_data[['pval']], HR_pancohort[[paste0('p-val_',testVAR)]])


pval_vec = vector("character", length(P_value_raw))
for (i in 1:length(P_value_raw)){
  pval = P_value_raw[i]
  if (pval>=0.1){
    pval_vec[i] = as.character(round(pval,2))
  }else{
    pval_vec[i] = format(pval, scientific = TRUE, digits = 2)  # Format for scientific notation
  }
}

plot_data$pval = pval_vec[1:length(pval_vec)-1]

xmin_lim = 0
xmax_lim = 1.5
breaks_x = c(0, 0.5, 1, 1.5)
labels_x = breaks_x


fontSize = 1.2

pdf_file <- paste0(result_dir,paste0("NSCLCcohorts_Forest_",survival_type,"_",HR_type,"_",Var_type,"_",LLRmodelNA,"_",testVAR0,".pdf"))
fig_width = 5
fig_height = 3
pdf(pdf_file, onefile=FALSE, width = fig_width,height=fig_height)


plot_data %>%
  forestplot(labeltext = c("dataset", "size", "pval"),
             graph.pos = 3,
             boxsize = 0.35,
             vertices = TRUE,
             clip = c(xmin_lim, xmax_lim),
             xlog = FALSE,
             zero = 1,
             txt_gp = fpTxtGp(ticks=gpar(cex=fontSize),xlab=gpar(cex=fontSize),label=gpar(cex=fontSize),legend=gpar(cex=fontSize),title=gpar(cex=fontSize)), # label, summary, xlab, title, ticks legend
             xlab = paste0("Hazard ratio (",testVAR_name,")"),
             xticks = breaks_x,
             graphwidth = unit(3, "cm"),
             lineheight = unit(2, "cm")
             ) %>%
  fp_set_style(box = "black",
               line = "black",
               summary = "black") %>%
  fp_add_header(dataset = c("Data set")|> fp_txt_plain() |> fp_align_center(),
                size = c("Samples")|> fp_txt_plain() |> fp_align_center(), # "No. of patients"
                pval = c("P-value")|> fp_txt_plain() |> fp_align_center())

dev.off()

```

